\documentclass{article}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\title{Exercise on Project Structure}

\begin{document}
\maketitle
\tableofcontents

\newpage
\section{First section}
Jelly beans tiramisu chocolate bar chupa chups candy canes lollipop liquorice cheesecake tootsie roll. Marshmallow sugar plum cake tiramisu jelly cake bear claw. Croissant liquorice jelly lollipop carrot cake. Cheesecake danish chocolate halvah. Candy canes jujubes chupa chups jujubes candy canes tootsie roll toffee danish sesame snaps. Sugar plum gummi bears lollipop chocolate bar. Ice cream soufflé sweet lemon drops chocolate soufflé gingerbread cookie muffin.

Sugar plum danish halvah jujubes pastry donut carrot cake dessert donut. Sweet gingerbread gingerbread wafer gummi bears pastry. Tootsie roll donut powder cake. Brownie halvah lollipop gummi bears. Liquorice pie caramels jelly beans. Halvah danish brownie danish toffee cotton candy lemon drops.

Donut jelly beans candy canes muffin lemon drops carrot cake sugar plum. Sesame snaps pie bonbon cotton candy cookie caramels. Chupa chups dessert tootsie roll carrot cake candy cookie sweet roll jelly beans. Oat cake halvah gingerbread bonbon sweet sesame snaps dragée carrot cake. Bonbon pastry chupa chups. Chocolate cake macaroon jelly-o. Pastry pastry sugar plum jujubes lemon drops gummi bears sugar plum cheesecake fruitcake. Chocolate cake chocolate candy. Wafer carrot cake chocolate cake jujubes cupcake soufflé gingerbread chocolate cake donut.

\newpage
\section{Second section}

Sweet roll tart chocolate bar cotton candy jelly beans cheesecake. Ice cream pastry halvah tart dragée jelly beans chocolate toffee wafer. Pie tootsie roll wafer halvah jelly cake donut cheesecake. Caramels tiramisu candy canes tootsie roll danish apple pie topping topping. Gingerbread carrot cake chocolate cake liquorice jujubes caramels cupcake apple pie macaroon.

Marzipan marzipan cupcake pastry tart sesame snaps sweet roll pie. Marshmallow pastry icing macaroon topping toffee powder gummi bears soufflé. Danish carrot cake sesame snaps cotton candy candy canes cotton candy. Cheesecake dragée sweet roll. Cheesecake muffin tiramisu marshmallow. Lollipop lollipop jujubes jelly-o cake sugar plum. Lemon drops donut topping wafer. Pastry gingerbread biscuit cheesecake muffin candy canes sweet pastry. 

Carrot cake cotton candy toffee jelly-o chocolate toffee cookie sweet roll. Pastry tootsie roll marzipan apple pie lollipop gingerbread sweet. Fruitcake gummies bear claw croissant. Pie cotton candy cheesecake powder bonbon candy canes lemon drops sweet.

\end{document}
